---
title: "Maps using VIF variables"
author: "Gabriel de Freitas Pereira"
date: "19/05/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(leaflet)
library(raster)
library(leafem)
library(rgdal)
```


## Suitability Map - Geographical Filtering WorldClim 2.1

AUC = 0.87. For more details see <https://bnutmodelling.shinyapps.io/vif_vars_geog/>.

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}

#importing data from the geographical models using geoglyphs and tp_avg_aggr:
presence_2_5 <- raster("./vif_vars_geog/geog 2.5 tp_avg_aggr and geoglyphs vif/ensembles/presence/Bertholletia excelsa_present.tif")
presence_2_5[presence_2_5 == 0] <- NA

suitability_2_5 <- raster("./vif_vars_geog/geog 2.5 tp_avg_aggr and geoglyphs vif/ensembles/suitability/Bertholletia excelsa_present.tif")
suitability_2_5 <- mask(suitability_2_5, presence_2_5)

#building colors gradient:
fun <- colorRampPalette(c("#FF5948","#FEFC00", "#2C7202"))
#determining the values range according with the raster used:
cc <- fun(413)
#building palette to be used into leaflet():
pal <- colorNumeric(palette = cc,values(suitability_2_5), na.color = "transparent")
#building the map
leaflet() %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(suitability_2_5, opacity = 0.6, 
                 colors = pal) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(suitability_2_5), 
            opacity = 1, title = "Presence suitability (558 - 971)") 


```

## Future Projections - SSP245
 
CMIP6 data. For more details click [here](https://bideco.shinyapps.io/futur3/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}

ssp245_6mods_ <- raster("./futur3/combined_maps_g/Bertholletia excelsa_ssp245_2041_2060_change_6mods.tif")
pal <- colorFactor(palette = c("red","green","blue"), domain = values(ssp245_6mods_), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable","Become suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(ssp245_6mods_, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(ssp245_6mods_), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 6mods - ssp245")

```

## Future Projections - SSP585

CMIP6 data. For more details click [here](https://bideco.shinyapps.io/futur3/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}

ssp585_6mods_ <- raster("./futur3/combined_maps_g/Bertholletia excelsa_ssp585_2041_2060_change_6mods.tif") 
pal <- colorFactor(palette = c("red","green"), domain = values(ssp585_6mods_), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(ssp585_6mods_, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(ssp585_6mods_), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 6mods - ssp585")


```





## Suitability Map - Geographical Filtering WorldClim 1.4 LAM

AUC = 0.88. For more details see <https://bideco.shinyapps.io/cmip5_prsent/>.

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}

#importing data from the geographical models using geoglyphs and tp_avg_aggr:
#presence points used for modelling / binary map 0 absence and 1 presence
presence_2_5 <- raster("./cmip5_prsent/geog 2.5 tp_avg_aggr and geoglyphs vif lam cmip5/ensembles/presence/Bertholletia excelsa_present.tif")
presence_2_5[presence_2_5 == 0] <- NA

#suitability map 0-1000:
suitability_2_5 <- raster("./cmip5_prsent/geog 2.5 tp_avg_aggr and geoglyphs vif lam cmip5/ensembles/suitability/Bertholletia excelsa_present.tif")
suitability_2_5 <- mask(suitability_2_5, presence_2_5)

fun <- colorRampPalette(c("#FF5948","#FEFC00", "#2C7202"))
#determining the values range according with the raster used:
cc <- fun(432)
#building palette to be used into leaflet():
pal <- colorNumeric(palette = cc,values(suitability_2_5), na.color = "transparent")
#building the map
leaflet() %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(suitability_2_5, opacity = 0.6, 
                 colors = pal) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(suitability_2_5), 
            opacity = 1, title = "Presence suitability (539 - 971)")


```

## Future Projections - RCP45
 
CMIP5 data. For more details click [here](https://bnutmodelling.shinyapps.io/future_cmip5/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
rcp45_3mods_ <- raster("./future_cmip5/combined_maps_g/Bertholletia excelsa_RCP45_2040_2069_change_3mods.tif")
pal <- colorFactor(palette = c("red","green","blue"), domain = values(rcp45_3mods_), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable","Become suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(rcp45_3mods_, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(rcp45_3mods_), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 3mods - rcp45")

```

## Future Projections - RCP85

CMIP5 data. For more details click [here](https://bnutmodelling.shinyapps.io/future_cmip5/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
rcp85_3mods_ <- raster("./future_cmip5/combined_maps_g/Bertholletia excelsa_RCP85_2040_2069_change_3mods.tif")
pal <- colorFactor(palette = c("red","green","blue"), domain = values(rcp85_3mods_), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable","Become suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(rcp85_3mods_, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(rcp85_3mods_), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 3mods - rcp85")


```








## Suitability Map - Environmental Filtering

AUC = 0.94. For more details see <https://bideco.shinyapps.io/vif_vars_env/>.

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
presence_2_5_ <- raster("./vif_vars_env/env 2.5 tp_avg_aggr and geoglyphs vif/ensembles/presence/Bertholletia excelsa_present.tif")
presence_2_5_[presence_2_5_ == 0] <- NA

#suitability map 0-1000:
suitability_2_5_ <- raster("./vif_vars_env/env 2.5 tp_avg_aggr and geoglyphs vif/ensembles/suitability/Bertholletia excelsa_present.tif")
suitability_2_5_ <- mask(suitability_2_5_, presence_2_5_)

fun <- colorRampPalette(c("#FF5948","#FEFC00", "#2C7202"))
            #determining the values range according with the raster used:
            cc <- fun(484)
            #building palette to be used into leaflet():
            pal <- colorNumeric(palette = cc,values(suitability_2_5_), na.color = "transparent")
            #building the map
            leaflet() %>%
                addProviderTiles(providers$OpenStreetMap) %>%
                setView(lng = -55.0, lat = -5, zoom = 4) %>%
                addRasterImage(suitability_2_5_, opacity = 0.6, 
                               colors = pal) %>% 
                addLegend(position = "bottomright", pal = pal, values = values(suitability_2_5_), 
                          opacity = 1, title = "Presence suitability (503 - 987)")

```

## Future Projections - SSP245
 
CMIP6 data. For more details click [here](https://bnutmodelling.shinyapps.io/future_env/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
ssp245_6mods <- raster("./future_env/combined_maps_g_without_lam/Bertholletia excelsa_ssp245_2041_2060_change_6mods.tif")

pal <- colorFactor(palette = c("red","green","blue"), domain = values(ssp245_6mods), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable","Become suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(ssp245_6mods, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(ssp245_6mods), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 6mods - ssp245")

```

## Future Projections - SSP585

CMIP6 data. For more details click [here](https://bnutmodelling.shinyapps.io/future_env/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
ssp585_6mods <- raster("./future_env/combined_maps_g_without_lam/Bertholletia excelsa_ssp585_2041_2060_change_6mods.tif")

#6 mods - 585
pal <- colorFactor(palette = c("red","green","blue"), domain = values(ssp585_6mods), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable","Become suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(ssp585_6mods, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(ssp585_6mods), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 6mods - ssp585")



```






## Suitability Map - Environmental Filtering LAM

AUC = 0.93. For more details see <https://bideco.shinyapps.io/vif_vars_env/>.

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
presence_2_5 <- raster("./vif_vars_env/env 2.5 tp_avg_aggr and geoglyphs vif lam/ensembles/presence/Bertholletia excelsa_present.tif")
presence_2_5[presence_2_5 == 0] <- NA

#suitability map 0-1000:
suitability_2_5 <- raster("./vif_vars_env/env 2.5 tp_avg_aggr and geoglyphs vif lam/ensembles/suitability/Bertholletia excelsa_present.tif")
suitability_2_5 <- mask(suitability_2_5, presence_2_5)


fun <- colorRampPalette(c("#FF5948","#FEFC00", "#2C7202"))
#determining the values range according with the raster used:
cc <- fun(392)
#building palette to be used into leaflet():
pal <- colorNumeric(palette = cc,values(suitability_2_5), na.color = "transparent")
#building the map
leaflet() %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(suitability_2_5, opacity = 0.6, 
                 colors = pal) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(suitability_2_5), 
            opacity = 1, title = "Presence suitability (588 - 980)")

```

## Future Projections - SSP245 LAM
 
CMIP6 data. For more details click [here](https://bnutmodelling.shinyapps.io/future_env/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
#env 2.5 vif considering geoglyphs and tp:
ssp245_6mods_ <- raster("./future_env/combined_maps_g/Bertholletia excelsa_ssp245_2041_2060_change_6mods.tif")

#6 mods - 245
pal <- colorFactor(palette = c("red","green","blue"), domain = values(ssp245_6mods_), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable","Become suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(ssp245_6mods_, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(ssp245_6mods_), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 6mods - ssp245")


```

## Future Projections - SSP585 LAM

CMIP6 data. For more details click [here](https://bnutmodelling.shinyapps.io/future_env/).

```{r, warning=FALSE, echo=FALSE, message=FALSE, eval=TRUE}
ssp585_6mods_ <- raster("./future_env/combined_maps_g/Bertholletia excelsa_ssp585_2041_2060_change_6mods.tif")

#6 mods - 585
pal <- colorFactor(palette = c("red","green","blue"), domain = values(ssp585_6mods_), na.color = "transparent")
labels <- c("Unsuitable","Remain suitable","Become suitable")

leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  setView(lng = -55.0, lat = -5, zoom = 4) %>%
  addRasterImage(ssp585_6mods_, opacity = 0.6, 
                 colors = pal, project = FALSE) %>% 
  addLegend(position = "bottomright", pal = pal, values = values(ssp585_6mods_), 
            opacity = 1, labFormat = function(type,cuts,p){paste0(labels)}, 
            title = "Presence suitability  <p> 6mods - ssp585")

```

